### 2020-09-28 meeting materials/thesis_meetingsept28.Rmd
meeting materials for sept 28th thesis meeting, including rmd summary document and importing extracted data

---
title: "Thesis"
author: "Maggie Slein"
date: "9/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Goals for today's meeting
1. Discuss figure(s) from preliminary extraction 
2. Outline extract methods and binning
3. Next steps for conducting hollistic search terms

###Study selection

Khelifa et al 2019

I really enjoy this paper and feel that its figures and methods are very straightforward. I'd like my final figure to ultimately look like this

```{r}
knitr::include_graphics("khelifa-fig2.png")
```

```{r, echo= FALSE}
library(tidyverse)
extract<-read_csv("prelim_extact.csv")
names(extract)

ggplot(extract, aes(x=org_level, y=cohen_d, color=fluctuation_pattern))+
  geom_boxplot()+
  theme_classic()+
  ggtitle("Cohen D values for effect of flux patterns across different levels of organization")

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.