### 2021-02-22 shared analysis files /data_wrangling_plot22feb.Rmd
updated plots

---
title: "data wrangling and plots"
author: "Maggie Slein"
date: "2/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}

## loading libraries 
pacman::p_load(metafor, tidyverse)
#devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
#library(orchaRd)

#load data
dat_MA_0<- read_csv("maggie_MA_2_8_21.csv")
```

```{r, echo= FALSE}
##wrangling calculations for metafor

#make standardized SD columns based on N and SE vs SD
dat_MA <-dat_MA_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_MA_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_MA, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% filter (!is.na(yi))

#fit initial MA model, abs (note structure )
mod1 <- rma.mv(abs(yi), vi, data=dat_MA_ES, 
                 method="REML", random = ~1 | response_id/ experiment_id/ study_id) 
summary(mod1)

#fit dirrectional model
mod2<-rma.mv(yi, vi, data=dat_MA_ES, 
             method="REML", random = ~1 | response_id/ experiment_id/ study_id) 
summary(mod2)
```

```{r, echo= FALSE}
## data wrangling for plots
#count and propotions for the different variables
count(dat_MA_ES, flux_range) #46 rows w/ range of 10, 51 w/ range of 15, 36 w/range of 6
count(dat_MA_ES, resp_units, sort = TRUE) 
count(dat_MA_ES, mean_temp_constant, study_id) #24 C common accross 5 studies
count(dat_MA_ES, resp_def, sort = TRUE) # body mass, avg. cum. num of eggs laid/fm, avg. num of eggs laid/fm, survival, rate of change, are the the most common reponses
count(dat_MA_ES, genus, study_id) # 3 studies featuring drosophila
count(dat_MA_ES, genus)# 16 genuses to bin by 
count(dat_MA_ES, size)# good spread for body size
count(dat_MA_ES, larger_group) # pretty much mostly animal studies

#removing outlier point to get better picture of data structure
normalized <- dat_MA_ES %>%
  filter(yi > -20) 
#filtering for studies with common variable C (CT assays)
common_unit_C <-dat_MA_ES %>% 
  filter(resp_units == "C")
#filtering for studies with common variable g 
common_unit_g <-dat_MA_ES %>% 
filter(resp_units == "g")
#filtering for genus to only be drosophila 
drosophila <- dat_MA_ES %>% 
  filter(genus == "Drosophila")
#filtering for just flux ranges of 10 C 
common_range <- dat_MA_ES %>% 
  filter(flux_range == "10")
#filtering for just studies with mean temp of 24
common_temp <- dat_MA_ES %>% 
  filter(mean_temp_constant == "24")
```

```{r, echo= FALSE}
#SMD across all studies
ggplot(normalized, aes(x=study_id, y=yi))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45))+
  ggtitle("SMD across all studies")
```

```{r, echo= FALSE}
# scatterplot of standardized mean response faceted by exp age
ggplot(dat_MA_ES, aes(x=log(vi), y=log(yi), color = study_id))+
  geom_point()+
  theme_bw()+
  facet_wrap(~exp_age)+
  ggtitle("Log transformed SMD across log transformed variance 
          colored by study and faceted by experimental age")
```

```{r, echo= FALSE}
# scatterplot of standardized mean response vs mean temp faceted by org level and colored by flux_range
ggplot(dat_MA_ES, aes(x=mean_temp_constant, y=yi, color = as.factor(flux_range)))+
  geom_point()+
  facet_wrap(~org_level)+
  theme_bw()+
  ggtitle("SMD across mean temperature colored by fluctation range 
          and faceted by organization level")
```

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  ggtitle("SMD across fluctuation ranges colored by organization level 
          and fit with linear model")
```

```{r, echo= FALSE}
ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_minimal()+
  ggtitle("SMD across fluctuation ranges colored by 
          studies and fit with linear model by study")
```

```{r, echo= FALSE}
#trying to look at drosophila response across studies 
ggplot(drosophila, aes(y=yi, x=study_id, color = study_id))+
  geom_boxplot()+
  theme_minimal()+
  ggtitle("SMD across studies with the same study genus (Drosophila)")
```

```{r, echo= FALSE}
ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
  geom_jitter(alpha = 0.5)+
  theme_minimal()+
  ggtitle("SMD across fluctuation ranges colored by genus")
```

```{r, echo= FALSE}
#looking at yi across studies with the same temperature range (10 C)
ggplot(common_range, aes(y=yi, x=study_id, color = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
                                  ggtitle("SMD across studies with the same 
                                          temperature fluctuation range (10 C)")
```

```{r, echo= FALSE}
#looking at yi across studies with the same mean temperature (24 C)
ggplot(common_temp, aes(y=yi, x=study_id, color = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 7, angle = 45))+
  ggtitle("SMD across studies with the same mean temperature (24 C)")
```

```{r, echo= FALSE}
#trying to look at how size trends yi across increasing flux ranges 
ggplot(normalized, aes(y=abs(yi), x=flux_range, color = as.factor(size)))+
  geom_point(alpha = 0.5)+
  geom_smooth(formula = y~x)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 7, angle = 45))+
  ggtitle("Absolute SMD across fluctuation ranges colored 
          by size of study organism and fitted with model curve")
```

```{r, echo= FALSE}
ggplot(common_unit_C, aes(y=yi, x=study_id, color = as.factor(flux_range)))+
  geom_point(alpha = 0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 7, angle = 45))+
  ggtitle("SMD across studies with the common unit C (CT assays)")
```

```{r, echo= FALSE}
#plot across studies with common unit gram
ggplot(common_unit_g, aes(y=yi, x=flux_range, color = study_id))+
  geom_point(alpha =0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
```