### 2021-03-21 thesis draft/thesis.Rmd
trying to make an r file for this thesis draft?

<!--
This is for including Chapter 1.  Notice that it's also good practice to name your chunk.  This will help you debug potential issues as you knit.  The chunk above is called intro and the one below is called chapter1.  Feel free to change the name of the Rmd file as you wish, but don't forget to change it here from chap1.Rmd.
-->

<!--
The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of Chapter 1.  These labels will automatically be generated (if not specified) by changing the spaces to hyphens and capital letters to lowercase.  Look for the reference to this label at the beginning of Chapter 2.
-->

# R Markdown Basics {#rmd-basics}

Here is a brief introduction into using _R Markdown_. _Markdown_ is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. _R Markdown_ provides the flexibility of _Markdown_ with the implementation of **R** input and output.  For more details on using _R Markdown_ see <https://rmarkdown.rstudio.com>.

Be careful with your spacing in _Markdown_ documents.  While whitespace largely is ignored, it does at times give _Markdown_ signals as to how to proceed.  As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph.  In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).

## Lists

It's easy to create a list.  It can be unordered like

* Item 1
* Item 2

or it can be ordered like

1. Item 1
4. Item 2

Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a tab).  (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
    - Item 3a
    - Item 3b

## Line breaks

Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
This should be a new paragraph.

*Now for the correct way:*

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.

This should be a new paragraph.

## R chunks

When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document. You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):

```{r cars}
summary(cars)
```

## Inline code

If you'd like to put the results of your analysis directly into your discussion, add inline code like this:

> The `cos` of $2 \pi$ is `r cos(2*pi)`.

Another example would be the direct calculation of the standard deviation:

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.

One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be indented.

As you see with `$2 \pi$` above, mathematics can be added by surrounding the mathematical text with dollar signs.  More examples of this are in [Mathematics and Science] if you uncomment the code in [Math].

## Including plots

You can also embed plots. For example, here is a way to use the base **R** graphics package to produce a plot using the built-in `pressure` dataset:

```{r pressure, echo=FALSE, cache=TRUE, fig.height=3, fig.width=5}
plot(pressure)
```

Note that the `echo=FALSE` parameter was added to the code chunk to prevent printing of the **R** code that generated the plot. There are plenty of other ways to add chunk options (like `fig.height` and `fig.width` in the chunk above).  More information is available at <https://yihui.org/knitr/options/>.

Another useful chunk option is the setting of `cache=TRUE` as you see here.  If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.  Later in this file, you'll see a way to reference plots created in **R** or external figures.

## Loading and exploring data

Included in this template is a file called `flights.csv`.  This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014. More information about this dataset and its **R** package is available at <https://github.com/ismayc/pnwflights14>. This subset includes only Portland flights and only rows that were complete with no missing values. Merges were also done with the `airports` and `airlines` data sets in the `pnwflights14` package to get more descriptive airport and airline names.

We can load in this data set using the following commands:

```{r load_data}
# flights.csv is in the data directory
flights_path <- here::here("data", "flights.csv")
# string columns will be read in as strings and not factors now
flights <- read.csv(flights_path, stringsAsFactors = FALSE)
```

The data is now stored in the data frame called `flights` in **R**.  To get a better feel for the variables included in this dataset we can use a variety of functions. Here we can see the dimensions (rows by columns) and also the names of the columns.

```{r str}
dim(flights)
names(flights)
```

Another good idea is to take a look at the dataset in table form.  With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here. I recommend you enter the command into the Console **_after_** you have run the **R** chunks above to load the data into **R**.

```{r view_flights, eval=FALSE}
View(flights)
```

While not required, it is highly recommended you use the `dplyr` package to manipulate and summarize your data set as needed.  It uses a syntax that is easy to understand using chaining operations.  Below I've created a few examples of using `dplyr` to get information about the Portland flights in 2014.  You will also see the use of the `ggplot2` package, which produces beautiful, high-quality academic visuals.

We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:

```{r load_pkgs, message=FALSE}
# List of packages required for this analysis
pkg <- c("dplyr", "ggplot2", "knitr", "bookdown")
# Check if packages are not installed and assign the
# names of the packages not installed to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]
# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg)) {
  install.packages(new.pkg, repos = "https://cran.rstudio.com")
}
# Load packages
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
```

\clearpage

The example we show here does the following:

- Selects only the `carrier_name` and `arr_delay` from the `flights` dataset and then assigns this subset to a new variable called `flights2`.

- Using `flights2`, we determine the largest arrival delay for each of the carriers.

```{r max_delays}
flights2 <- flights %>%
  select(carrier_name, arr_delay)
max_delays <- flights2 %>%
  group_by(carrier_name) %>%
  summarize(max_arr_delay = max(arr_delay, na.rm = TRUE))
```

A useful function in the `knitr` package for making nice tables in _R Markdown_ is called `kable`.  It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.  This again goes to show how nice reproducible documents can be! (Note the use of `results="asis"`, which will produce the table instead of the code to create the table.)  The `caption.short` argument is used to include a shorter title to appear in the List of Tables.

```{r maxdelays, results="asis"}
kable(max_delays,
  col.names = c("Airline", "Max Arrival Delay"),
  caption = "Maximum Delays by Airline",
  caption.short = "Max Delays by Airline",
  longtable = TRUE,
  booktabs = TRUE
)
```

The last two options make the table a little easier-to-read.

We can further look into the properties of the largest value here for American Airlines Inc.  To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original `flights` dataset.

```{r max_props}
flights %>%
  filter(
    arr_delay == 1539,
    carrier_name == "American Airlines Inc."
  ) %>%
  select(-c(
    month, day, carrier, dest_name, hour,
    minute, carrier_name, arr_delay
  ))
```

We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth.  Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.

```{r march3plot, fig.height=3, fig.width=6}
flights %>%
  filter(month == 3, day == 3) %>%
  ggplot(aes(x = dep_time, y = arr_delay)) +
  geom_point()
```

## Additional resources

- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>

- _R Markdown_
    - Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf>

- _RStudio IDE_
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf>
    - Official website - <https://rstudio.com/products/rstudio/>

- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html>

- `ggplot2`
    - Documentation - <https://ggplot2.tidyverse.org/>
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf>

<!--chapter:end:01-chap1.Rmd-->

# Mathematics and Science {#math-sci}

<!-- Required to number equations in HTML files -->
<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { equationNumbers: { autoNumber: "AMS" } }
});
</script>

## Math

\TeX\ is the best way to typeset mathematics. Donald Knuth designed \TeX\ when he got frustrated at how long it was taking the typesetters to finish his book, which contained a lot of mathematics.  One nice feature of _R Markdown_ is its ability to read LaTeX code directly.

If you are doing a thesis that will involve lots of math, you will want to read the following section which has been commented out. If you're not going to use math, skip over or delete this next commented section.

<!-- MATH and PHYSICS majors: Uncomment the following section -->
<!--
$$\sum_{j=1}^n (\delta\theta_j)^2 \leq {{\beta_i^2}\over{\delta_i^2 + \rho_i^2}}
\left[ 2\rho_i^2 + {\delta_i^2\beta_i^2\over{\delta_i^2 + \rho_i^2}} \right] \equiv \omega_i^2
$$

From Informational Dynamics, we have the following (Dave Braden):

After _n_ such encounters the posterior density for $\theta$ is

$$
\pi(\theta|X_1< y_1,\dots,X_n<y_n) \varpropto \pi(\theta) \prod_{i=1}^n\int_{-\infty}^{y_i}
   \exp\left(-{(x-\theta)^2\over{2\sigma^2}}\right)\ dx
$$

Another equation:

$$\det\left|\,\begin{matrix}%
c_0&c_1\hfill&c_2\hfill&\ldots&c_n\hfill\cr
c_1&c_2\hfill&c_3\hfill&\ldots&c_{n+1}\hfill\cr
c_2&c_3\hfill&c_4\hfill&\ldots&c_{n+2}\hfill\cr
\,\vdots\hfill&\,\vdots\hfill&
  \,\vdots\hfill&&\,\vdots\hfill\cr
c_n&c_{n+1}\hfill&c_{n+2}\hfill&\ldots&c_{2n}\hfill\cr
\end{matrix}\right|>0$$

Lapidus and Pindar, Numerical Solution of Partial Differential Equations in Science and
Engineering.  Page 54

$$
\int_t\left\{\sum_{j=1}^3 T_j \left({d\phi_j\over dt}+k\phi_j\right)-kT_e\right\}w_i(t)\ dt=0,
   \qquad\quad i=1,2,3.
$$

L\&P  Galerkin method weighting functions.  Page 55

$$
\sum_{j=1}^3 T_j\int_0^1\left\{{d\phi_j\over dt} + k\phi_j\right\} \phi_i\ dt
   = \int_{0}^1k\,T_e\phi_idt, \qquad i=1,2,3 $$

Another L\&P (p145)

$$
\int_{-1}^1\!\int_{-1}^1\!\int_{-1}^1 f\big(\xi,\eta,\zeta\big)
   = \sum_{k=1}^n\sum_{j=1}^n\sum_{i=1}^n w_i w_j w_k f\big( \xi,\eta,\zeta\big).
$$

Another L\&P (p126)

$$
\int_{A_e} (\,\cdot\,) dx dy = \int_{-1}^1\!\int_{-1}^1 (\,\cdot\,) \det[J] d\xi d\eta.
$$
-->

## Chemistry 101: Symbols

Chemical formulas will look best if they are not italicized. Get around math mode's automatic italicizing in LaTeX by using the argument `$\mathrm{formula here}$`, with your formula inside the curly brackets.  (Notice the use of the backticks here which enclose text that acts as code.)

So, $\mathrm{Fe_2^{2+}Cr_2O_4}$ is written `$\mathrm{Fe_2^{2+}Cr_2O_4}$`.

<!--
The \noindent command below does what you'd expect:  it forces the current line/paragraph to not indent. This was done here to match the format of the LaTeX thesis PDF.
-->

\noindent Exponent or Superscript: $\mathrm{O^-}$

\noindent Subscript: $\mathrm{CH_4}$

To stack numbers or letters as in $\mathrm{Fe_2^{2+}}$, the subscript is defined first, and then the superscript is defined.

\noindent Bullet: CuCl $\bullet$ $\mathrm{7H_{2}O}$

\noindent Delta: $\Delta$

\noindent Reaction Arrows: $\longrightarrow$ or  $\xrightarrow{solution}$

\noindent Resonance Arrows: $\leftrightarrow$

\noindent Reversible Reaction Arrows: $\rightleftharpoons$

### Typesetting reactions

You may wish to put your reaction in an equation environment, which means that LaTeX will place the reaction where it fits and will number the equations for you.

\begin{equation}
  \mathrm{C_6H_{12}O_6  + 6O_2} \longrightarrow \mathrm{6CO_2 + 6H_2O}
  (\#eq:reaction)
\end{equation}

We can reference this combustion of glucose reaction via Equation \@ref(eq:reaction).

### Other examples of reactions

$\mathrm{NH_4Cl_{(s)}}$ $\rightleftharpoons$ $\mathrm{NH_{3(g)}+HCl_{(g)}}$

\noindent $\mathrm{MeCH_2Br + Mg}$ $\xrightarrow[below]{above}$ $\mathrm{MeCH_2\bullet Mg \bullet Br}$

## Physics

Many of the symbols you will need can be found on the math page <https://web.reed.edu/cis/help/latex/math.html> and the Comprehensive LaTeX Symbol Guide (<https://mirror.utexas.edu/ctan/info/symbols/comprehensive/symbols-letter.pdf>).

## Biology

You will probably find the resources at <https://www.lecb.ncifcrf.gov/~toms/latex.html> helpful, particularly the links to bsts for various journals. You may also be interested in TeXShade for nucleotide typesetting (<https://homepages.uni-tuebingen.de/beitz/txe.html>).  Be sure to read the proceeding chapter on graphics and tables.

<!--chapter:end:02-chap2.Rmd-->

```{r include_packages_2, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis and also two functions
# used for labeling and referencing
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("remotes")',
        "first in the Console."
      )
    )
  }
}
if (!require(dplyr)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("dplyr", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("dplyr")',
        "first in the Console."
      )
    )
  }
}
if (!require(ggplot2)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("ggplot2", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("ggplot2")',
        "first in the Console."
      )
    )
  }
}
if (!require(bookdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("bookdown", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("bookdown")',
        "first in the Console."
      )
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
flights <- read.csv("data/flights.csv", stringsAsFactors = FALSE)
```

# Graphics, References, and Labels {#ref-labels}

## Figures

If your thesis has a lot of figures, _R Markdown_ might behave better for you than that other word processor.  One perk is that it will automatically number the figures accordingly in each chapter.    You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier.  If you label your figures, you can move the figures around and _R Markdown_ will automatically adjust the numbering for you.  No need for you to remember!  So that you don't have to get too far into LaTeX to do this, a couple **R** functions have been created for you to assist.  You'll see their use below.

<!--
One thing that may be annoying is the way _R Markdown_ handles "floats" like tables and figures (it's really \LaTeX's fault). \LaTeX\ will try to find the best place to put your object based on the text around it and until you're really, truly done writing you should just leave it where it lies. There are some optional arguments specified in the options parameter of the `label` function.  If you need to shift your figure around, it might be good to look here on tweaking the options argument:  <https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions>

If you need a graphic or tabular material to be part of the text, you can just put it inline. If you need it to appear in the list of figures or tables, it should be placed in a code chunk.
-->

In the **R** chunk below, we will load in a picture stored as `reed.jpg` in our main directory.  We then give it the caption of "Reed logo", the label of "reedlogo", and specify that this is a figure.  Make note of the different **R** chunk options that are given in the R Markdown file (not shown in the knitted document).

```{r reedlogo, fig.cap="Reed logo", out.width="0.2\\linewidth", fig.align="center"}
include_graphics(path = "figure/reed.jpg")
```

<!-- Note the use of `out.width` as a chunk option here. The resulting
image is 20% of what the linewidth is in LaTeX. You can also center
the image using `fig.align="center"` as shown.-->

Here is a reference to the Reed logo: Figure \@ref(fig:reedlogo).  Note the use of the `fig:` code here.  By naming the **R** chunk that contains the figure, we can then reference that figure later as done in the first sentence here.  We can also specify the caption for the figure via the R chunk option `fig.cap`.

\clearpage

<!-- clearpage ends the page, and also dumps out all floats.
  Floats are things like tables and figures. -->

Below we will investigate how to save the output of an **R** plot and label it in a way similar to that done above.  Recall the `flights` dataset from Chapter \@ref(rmd-basics).  (Note that we've shown a different way to reference a section or chapter here.)  We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.

```{r delaysboxplot, warnings=FALSE, messages=FALSE, fig.cap="Mean Delays by Airline", fig.width=6, fig.height=5}
mean_delay_by_carrier <- flights %>%
  group_by(carrier) %>%
  summarize(mean_dep_delay = mean(dep_delay))
ggplot(mean_delay_by_carrier, aes(x = carrier, y = mean_dep_delay)) +
  geom_bar(position = "identity", stat = "identity", fill = "red")
```

Here is a reference to this image: Figure \@ref(fig:delaysboxplot).

A table linking these carrier codes to airline names is available at <https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv>.

\clearpage

Next, we will explore the use of the `out.extra` chunk option, which can be used to shrink or expand an image loaded from a file by specifying `"scale= "`. Here we use the mathematical graph stored in the "subdivision.pdf" file.

```{r subd, results="asis", echo=FALSE, fig.cap="Subdiv. graph", out.extra="scale=0.75"}
include_graphics("figure/subdivision.pdf")
```

Here is a reference to this image: Figure \@ref(fig:subd).  Note that `echo=FALSE` is specified so that the **R** code is hidden in the document.

**More Figure Stuff**

Lastly, we will explore how to rotate and enlarge figures using the `out.extra` chunk option.  (Currently this only works in the PDF version of the book.)

```{r subd2, results="asis", echo=FALSE, out.extra="angle=180, scale=1.1", fig.cap="A Larger Figure, Flipped Upside Down"}
include_graphics("figure/subdivision.pdf")
```

As another example, here is a reference: Figure \@ref(fig:subd2).

## Footnotes and Endnotes

You might want to footnote something. ^[footnote text] The footnote will be in a smaller font and placed appropriately. Endnotes work in much the same way. More information can be found about both on the CUS site or feel free to reach out to <data@reed.edu>.

## Bibliographies

Of course you will need to cite things, and you will probably accumulate an armful of sources. There are a variety of tools available for creating a bibliography database (stored with the .bib extension).  In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero.  The Reed librarians have created Zotero documentation at <https://libguides.reed.edu/citation/zotero>.  In addition, a tutorial is available from Middlebury College at <https://sites.middlebury.edu/zoteromiddlebury/>.

_R Markdown_ uses _pandoc_ (<https://pandoc.org/>) to build its bibliographies.  One nice caveat of this is that you won't have to do a second compile to load in references as standard LaTeX requires. To cite references in your thesis (after creating your bibliography database), place the reference name inside square brackets and precede it by the "at" symbol.  For example, here's a reference to a book about worrying:  [@Molina1994].  This `Molina1994` entry appears in a file called `thesis.bib` in the `bib` folder.  This bibliography database file was created by a program called BibTeX.  You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the `bib` folder.

For more information about BibTeX and bibliographies, see our CUS site (<https://web.reed.edu/cis/help/latex/index.html>)^[@reedweb2007]. There are three pages on this topic:  _bibtex_ (which talks about using BibTeX, at <https://web.reed.edu/cis/help/latex/bibtex.html>), _bibtexstyles_ (about how to find and use the bibliography style that best suits your needs, at <https://web.reed.edu/cis/help/latex/bibtexstyles.html>) and _bibman_ (which covers how to make and maintain a bibliography by hand, without BibTeX, at <https://web.reed.edu/cis/help/latex/bibman.html>). The last page will not be useful unless you have only a few sources.

If you look at the YAML header at the top of the main .Rmd file you can see that we can specify the style of the bibliography by referencing the appropriate csl file.  You can download a variety of different style files at <https://www.zotero.org/styles>.  Make sure to download the file into the csl folder.

<!-- Fill the rest of the page with the content below for the PDF version. -->

\vfill

**Tips for Bibliographies**

- Like with thesis formatting, the sooner you start compiling your bibliography for something as large as thesis, the better. Typing in source after source is mind-numbing enough; do you really want to do it for hours on end in late April? Think of it as procrastination.
- The cite key (a citation's label) needs to be unique from the other entries.
- When you have more than one author or editor, you need to separate each author's name by the word "and" e.g. `Author = {Noble, Sam and Youngberg, Jessica},`.
- Bibliographies made using BibTeX (whether manually or using a manager) accept LaTeX markup, so you can italicize and add symbols as necessary.
- To force capitalization in an article title or where all lowercase is generally used, bracket the capital letter in curly braces.
- You can add a Reed Thesis citation^[@noble2002] option. The best way to do this is to use the phdthesis type of citation, and use the optional "type" field to enter "Reed thesis" or "Undergraduate thesis."

## Anything else?

If you'd like to see examples of other things in this template, please contact the Data @ Reed team (email <data@reed.edu>) with your suggestions. We love to see people using _R Markdown_ for their theses, and are happy to help.

<!--chapter:end:03-chap3.Rmd-->

# Conclusion {-}

If we don't want Conclusion to have a chapter number next to it, we can add the `{-}` attribute.

**More info**

And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph. Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.

<!--chapter:end:04-conclusion.Rmd-->

`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

# The First Appendix

<!--
If you feel it necessary to include an appendix, it goes here.
-->

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.

**In the main Rmd file**

```{r ref.label='include_packages', results='hide', echo = TRUE}
```

**In Chapter \@ref(ref-labels):**

```{r ref.label='include_packages_2', results='hide', echo = TRUE}
```

# The Second Appendix, for Fun

<!--chapter:end:05-appendix.Rmd-->

<!--
The bib chunk below must go last in this document according to how R Markdown renders.  More info is at http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
-->

\backmatter

<!-- 
If you'd like to change the name of the bibliography to something else,
delete "References" and replace it.
-->

# References {-}
<!--
This manually sets the header for this unnumbered chapter.
-->
\markboth{References}{References}
<!--
To remove the indentation of the first entry.
-->
\noindent

<!--
To create a hanging indent and spacing between entries.  These three lines may need to be removed for styles that don't require the hanging indent.
-->

\setlength{\parindent}{-0.20in}
\setlength{\leftskip}{0.20in}
\setlength{\parskip}{8pt}

<!--
This is just for testing with more citations for the bibliography at the end.  Add other entries into the list here if you'd like them to appear in the bibliography even if they weren't explicitly cited in the document.
-->

---
nocite: | 
  @angel2000, @angel2001, @angel2002a
...

<!--chapter:end:99-references.Rmd-->

---
title: 'My Final College Paper'
author: 'Margaret Anne Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
# If you have more two advisors, un-silence line 7
#altadvisor: 'Your Other Advisor'
department: 'Biology'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True

# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
  thesisdown::thesis_pdf: default 
#  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  I want to thank a few people.
dedication: |
  You can have a dedication here if you wish. 
preface: |
  This is an example of a thesis setup to use the reed thesis document class 
  (for LaTeX) and the R bookdown package, in general.

# Specify the location of the bibliography below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line).

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

Welcome to the _R Markdown_ thesis template. This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before.  Using _R Markdown_ will also allow you to easily keep track of your analyses in **R** chunks of code, with the resulting plots and output included as well.  The hope is this _R Markdown_ template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.

Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis.  The use of LaTeX in combination with _Markdown_ is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file. While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities.  After working with _Markdown_ and **R** together for a few weeks, we are confident this will be your reporting style of choice going forward.

<!-- 
If you're still on the fence about using _R Markdown_, check out the resource
for newbies available at <https://rbasics.netlify.com>. 
-->

**Why use it?**

_R Markdown_ creates a simple and straightforward way to interface with the beauty of LaTeX.  Packages have been written in **R** to work directly with LaTeX to produce nicely formatting tables and paragraphs. In addition to creating a user friendly interface to LaTeX, _R Markdown_ also allows you to read in your data, to analyze it and to visualize it using **R** functions, and also to provide the documentation and commentary on the results of your project.  Further, it allows for **R** results to be passed inline to the commentary of your results.  You'll see more on this later.

<!-- 
Having your code and commentary all together in one place has a plethora of 
benefits!
-->

**Who should use it?**

Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use _R Markdown_. Of particular use should be anyone in the sciences, but the user-friendly nature of _Markdown_ and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc. should make it of great benefit to nearly anyone writing a thesis project.

**For additional help with bookdown**

Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).

<!--chapter:end:thesis-draft.Rmd-->