### 2021-04-08 shared analysis files /data_wrangling_plot.Rmd
updates for plots for inclusion in results and discussion

---
title: "final figures for inclusion"
author: "Maggie Slein"
date: "4/7/21"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, results = 'hide', message=FALSE}
## loading libraries 
pacman::p_load(metafor, tidyverse, viridis, ggsci, plotly,  mgcv)
#devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
#library(orchaRd)

#load data
dat_MA_0<- read_csv("metafor.csv") %>%
  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
##wrangling calculations for metafor

#make standardized SD columns based on N and SE vs SD
dat_MA <-dat_MA_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_MA_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_MA, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) 
count(dat_MA_ES, study_id)

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -20) 
#filtering for just flux ranges of 10 C 
common_range <- dat_MA_ES %>% 
  filter(flux_range %in% c(10))
#filtering for just studies with mean temp of 24
common_temp <- dat_MA_ES %>% 
  filter(mean_temp_constant == "24")
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
## data wrangling for plots
#count and propotions for the different variables
count(dat_MA_ES, flux_range) #46 rows w/ range of 10, 51 w/ range of 15, 36 w/range of 6
count(dat_MA_ES, resp_units, sort = TRUE) 
count(dat_MA_ES, mean_temp_constant, study_id) #24 C common accross 5 studies
count(dat_MA_ES, resp_def, sort = TRUE) # body mass, avg. cum. num of eggs laid/fm, avg. num of eggs laid/fm, survival, rate of change, are the the most common reponses
count(dat_MA_ES, genus, study_id) # 3 studies featuring drosophila
count(dat_MA_ES, genus, species)# 16 genuses to bin by 
count(dat_MA_ES, size)# good spread for body size
count(dat_MA_ES, larger_group) # pretty much mostly animal studies
```

## Question: How does response compare across studies and experiments?
Hypothesis: There should be differences across studies because of differences in experimental designs and study organism that would mean different magnitudes of response.

Initial conclusions: The studies overall are different in their responses but not owing to study_id or experiment. However, the mixed effects model does suggest some differences when you include study _id as a mod.

```{r, echo= FALSE, message=FALSE}
#simple random effects model 
simple_model <- rma.mv(yi, vi, data=dat_MA_ES, 
               random = ~1 | study_id / experiment_id/ response_id,
                 method="REML") 
simple_model
```

## Relevant Plots

```{r, echo = FALSE}
#forest plot
forest <- normalized %>%
  group_by(study_id, experiment_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2")
```
**Figure 1.**
Forest plot broken out by study and experiment and their observed effected sizes and sampling variances.

```{r, echo = FALSE}
#funnel plot 
### set up 2x2 array for plotting
par(mfrow=c(2,2))
funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
```

```{r, echo = FALSE, fig.height=8, fig.width=12}
### draw funnel plots
#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
```

**Figure 2.** 
Well, from what I can gather, a majority of the effect sizes from our meta-analysis have a non-signficant effect size value. However, there is a population of a couple influential effect sizes that are significantly important.

```{r, echo= FALSE, message= FALSE}
#GAM APPROACH
#fit GAM
gam1<-gam(yi~te(mean_temp_constant,flux_range,k=7),data=normalized) # note, lots of decisions that I blew by here
summary(gam1)
#estimate values via gam
vals<-expand.grid(flux_range=seq(0,20,length.out = 100), mean_temp_constant=seq(0,35,length.out = 100))
pd.gam<-predict(gam1,newdata=vals)
surf<-data.frame(vals,value=pd.gam)
vertices <- data.frame(
  x = c(0, 0, 20),
  y = c(10, 0, 0), 
  id = c("a", "a", "a")
)
#plot values as a surface
```

```{r, echo= FALSE, message= FALSE}
#plot gam surface
ggplot()+
  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range, fill=value))+
  geom_contour(data = surf, 
               aes(x=mean_temp_constant,y=flux_range,z=value), colour=gray(0.2),binwidth=0.5,size=0.2)+
  geom_contour(data = surf,  aes(x=mean_temp_constant,y=flux_range,z=value), 
               colour="black", breaks=c(0),size=2)+
  #geom_contour_filled()
  scale_x_continuous('Mean Temp (°C)',expand=c(0,0))+
  scale_y_continuous('Temp Range (°C)',expand=c(0,0))+
  scale_fill_gradientn('Effect Size',
                       colors= viridis(100),
                       breaks=seq(-2,5,2),limits=c(-2,5),
                       guide = guide_colourbar(ticks.colour = gray(.2),
                                               ticks.linewidth = 1.,
                                               frame.colour = "black"))+
  coord_cartesian(xlim=c(0,35),ylim=c(0,20))+
  geom_jitter(data = normalized, mapping = aes(x=mean_temp_constant,y=flux_range,z= yi,fill = yi),
              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")+
  theme_bw()+ 
  theme(legend.position="bottom")+
  geom_polygon(data = vertices, mapping = aes(x = x, y = y), fill = "gray")
```
**Figure 3.** Contour plot of GAM generated effects sizes overlayed with raw effect sizes from the extracted dataset. Black lines denote the boundary from negative to positive effect sizes. Grey triange denotes excluded contour that lacks any raw data.

\newpage
## Question: How does response vary with study covariates?

Hypothesis: SMD between constant and fluctuating environments is most affected by temperature parameters (range and mean) but is also likely affected by demographic parameters (age, size, organization level)

Initial conclusions: Flux_range and mean temperature appear to be the most important contributers to variation in yi, though organization also contributes.

```{r}
#with interaction term 
full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML") 
full_rf_model

#separate model with thermal history trait as a mod 
full_rf_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = 
                          ~flux_range * mean_temp_constant + stressful, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
full_rf_thermal
#absolute yi model 
absolute_model<-rma.mv(abs(yi), vi, data=dat_MA_ES, mods = 
                          ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
absolute_model
```

## Relevant plots

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(org_level)))+
  geom_point()+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("SMD across fluctuation ranges colored by organization level 
          and fit with linear model")
```
**Figure 4.**
We can also see that there a difference in effect size in organization level responses across fluctuation range. Population level reponses trended positively across fluctuation ranges, while organism level responses trended negatively across fluctuation ranges.

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
  drop_na(stressful) %>%
ggplot(aes(x=flux_range, y=yi, color = factor(stressful)))+
  geom_jitter()+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("SMD across fluctuation ranges colored by 
          whether or mean temperature was thermally stressful")
```
**Figure 5.** 
When we consider where the temperatures in the fluctuation range reach thermally stressful levels, there is a small difference between responses to stressful  temperatures (negative) and non-stressful temperatures (positive).

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
ggplot(aes(x=flux_range, y=yi, color = resp_type))+
  geom_jitter()+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("SMD across fluctuation ranges colored by response type")

```
**Figure 6.** We cab see that the ranges at which traits are measured is much wider than rates, however, there appears to a minorly negative trend in effect sizes as traits increase in fluctuation range. The opposite is true for rates. Great way to show that there is a lack of information on rates in the collected dataset.

\newpage

Running questions:

- How do all the different models sound? Does it make sense to have these different models?
- Do the figures match up with the story/questions associated with them?
- Jittering versus standard plotting for the regression plots?
_________________________________
Supplementary Plots/Code

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#scatterplot of standardized mean difference across flux range colored by study fit with lm 
ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  ggtitle("SMD across fluctuation ranges colored by 
          studies and fit with linear model by study")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
# boxplots of how fluctuation range influences SMD
ggplot(normalized, aes(x=flux_range, y=yi))+
  geom_point(alpha = 0.5)+
  theme_bw()+
  geom_smooth(method="lm", formula = y~x)+
  ggtitle("SMD across flux_range")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#without interaction term
full_rf_model2<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + mean_temp_constant +
                        exp_age + size + org_level, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
full_rf_model2
#full model split out by stocks/rates 
full_rf_stocks<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
full_rf_stocks
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
# scatterplot of how mean temperature influences SMD
ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
  geom_point()+
  theme_minimal()+
  geom_smooth(formula = y~x, method = "lm")+
  ggtitle("SMD across mean temperature")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#SMD across all studies
ggplot(normalized, aes(x=reorder(study_id, -yi), y=yi, color = as.factor(experiment_id)))+
  geom_boxplot()+
  scale_color_tron()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45,hjust = 1), 
        legend.position = "bottom")+
  ggtitle("SMD across all studies")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#looking at yi across studies with the same temperature range (10 C)
#excluded wang because there were too few points to do a boxplot
common_range %>%
  filter(study_id != "wang2015") %>%
ggplot(aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
                                  ggtitle("SMD across studies with the same 
                                          temperature fluctuation range (10 C)")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
fig <- plot_ly(
  x = normalized$mean_temp_constant, 
  y = normalized$flux_range, 
  z = normalized$yi, 
  type = "contour", 
  contours = list(
    start = -19.42062,
    end = 5.286697,
    size = 1, 
    coloring = 'heatmap'
))
fig

```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#trying out mixed effects model 
fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
                 method="FE")
fig1me
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#boxplot of SMD across levels of organization
ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
  geom_boxplot(alpha =0.7)+
  geom_point(alpha = 0.3)+
  scale_fill_tron()+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggtitle("SMD across organization level")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
# scatterplot of standardized mean response faceted by exp age
ggplot(normalized, aes(x=yi, y=vi, color = study_id))+
  geom_point(alpha = 0.5)+
  theme_bw()+
  ggtitle("SMD across variance 
          colored by study")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}

#looking at yi across studies with the same mean temperature (24 C)
ggplot(common_temp, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
  geom_point()+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 7, angle = 45))+
  ggtitle("SMD across studies with the same mean temperature (24 C)")
#random effects model within subset data
fig13 <- rma.mv(yi, vi, data=common_temp,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig13
#simple linear model looking at how flux_range affects yi within subset data
simple13<-lm(yi~mean_temp_constant, data =common_temp)
summary(simple13)
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#boxplots of SMD across different genuses colored by body size
ggplot(normalized, aes(y=yi, x= reorder(genus, -yi), color = as.factor(size)))+
  geom_point()+
  theme_minimal()+
  ggtitle("SMD across genuses colored by body size")+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45,hjust = 1), 
        legend.position = "bottom")
#random effects model including genus as a random variable
fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~size,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig11
#simple linear model looking at how genus affects yi
simple11<-lm(yi~genus, data =dat_MA_ES)
summary(simple11)
```

```{r, echo = FALSE, eval= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#trying to look at drosophila response across studies 
ggplot(drosophila, aes(y=yi, x=as.factor(flux_range), color = study_id))+
  geom_boxplot()+
  geom_point()+
  scale_color_tron()+
  theme_bw()+
  ggtitle("SMD across studies with the same study genus (Drosophila)")
#random effects model including org_level as a random variable
fig10 <- rma.mv(yi, vi, data=drosophila, mods = ~flux_range + study_id,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig10
#simple linear model looking at how flux_range and study_id affect yi in drosphila specific studies
simple10<-lm(yi~flux_range + study_id, data =drosophila)
summary(simple10)
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
#random effects model including org_level as a random variable
fig6 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~mean_temp_constant,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig6
#simple linear model looking at how mean_temp_constant affects yi
simple6<-lm(yi~mean_temp_constant, data =dat_MA_ES)
summary(simple6)
```

```{r, echo= FALSE, results = 'hide', message=FALSE, eval = FALSE}
#basic linear model for study_id
simple1<-lm(yi ~ study_id, data = dat_MA_ES)
summary(fig1)
#basic linear model looking at how vi interacts with yi
simple2<-lm(yi ~ vi, data = dat_MA_ES)
summary(simple2)
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#boxplot of experimental ages and how their yi's stack up 
ggplot(normalized, aes(x=exp_age, y=yi, color = as.factor(exp_age)))+
  geom_boxplot()+
  theme_bw()+
  ggtitle("SMD across different experimental ages")

#random effects model including exp_age as a random variable
fig3 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~exp_age, 
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig3
#simple linear model looking at how exp_age affects yi
simple3<-lm(yi~exp_age, data =dat_MA_ES)
summary(simple3)
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#boxplot of different experimental ages and how they vi's stack up 
ggplot(normalized, aes(x=exp_age, y=vi, color = as.factor(exp_age)))+
  geom_boxplot()+
  theme_bw()+
  ggtitle("Variance across different experimental ages")

count(normalized, exp_age)

#simple linear model looking at how exp_age affects vi
simple4<-lm(vi~ exp_age, data =dat_MA_ES)
summary(simple4)
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
#random effects model including org_level as a random variable
fig5 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~org_level,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig5
#simple linear model looking at how org_level affects yi
simple5<-lm(yi~org_level, data =dat_MA_ES)
summary(simple5)
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
#full random effects model looking at all the responses in the entire dataset
response_variable_mod <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~resp_def, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
response_variable_mod

#full random effects model looking at all the response units in the entire dataset
response_units_mod <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~resp_units, 
               random = ~1 | experiment_id/ study_id /response_id,
                 method="REML") 
response_units_mod
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
#random effects model within subset data
fig12 <- rma.mv(yi, vi, data=common_range,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig12
#simple linear model looking at how flux_range affects yi within subset data
simple12<-lm(yi~flux_range, data =common_range)
summary(simple12)
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
#random effects model including org_level as a random variable
fig9 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + study_id,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig9
#simple linear model looking at how flux_range and study_id affect yi
simple9<-lm(yi~flux_range*study_id, data =dat_MA_ES)
summary(simple9)
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
#random effects model including org_level as a random variable
fig8 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + org_level,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig8
#simple linear model looking at how flux_range and org_level affect yi
simple8<-lm(yi~flux_range*org_level, data =dat_MA_ES)
summary(simple8)
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
#random effects model including org_level as a random variable
fig7 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig7 
#simple linear model looking at how flux_range affects yi
simple7<-lm(yi~flux_range, data =dat_MA_ES)
summary(simple7)
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
rf_model<-rma.mv(abs(yi), vi, data=dat_MA_ES, mods = ~flux_range +
                        exp_age + size + org_level + mean_temp_constant, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML")
rf_model
```

```{r, echo = FALSE, eval= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#plot across studies with common response variable body mass
ggplot(common_unit_g, aes(y=yi, color=as.factor(flux_range), x = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
#random effects model within subset data
fig14 <- rma.mv(yi, vi, data=common_unit_g,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig14
#simple linear model looking at how flux_range affects yi within subset data
simple14<-lm(yi~study_id, data =common_unit_g)
summary(simple14)
```