### 2021-04-28 writing /thesis_draft/index.Rmd

---
title: 'Mean temperature, fluctuation range, and biological organization affect biological responses'
author: 'Margaret A. Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
department: 'Biology'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site
params:
  'Install needed packages for {thesisdown}': True
output:
  thesisdown::thesis_pdf: default 
abstract: 'Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns. Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization. Here, we present a quantitative meta analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization. Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model. Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.'
acknowledgements: |

  The future is bright, the future is dark

  Time stands still, yet flies like a lark

  Memories endure and memories fade

  People are rough, yet smooth like suede

  The days seem long, the days seem short

  It rains all year, until the sunshine retorts

  Reed is an experience, Reed is a place

  Reed is a journey of learning how to keep pace

  Reed is a individual, Reed is a kollective

  Reed is space for being reflective

  Reed is the present, Reed is the past

  Reed is the net for our future that we cast

  To those from Reed

  To those from before

  To those from the outside

  To those at the core

  To those I still talk to

  To those I talk to no more

  To those still here today

  To those who now soar

  A few words of dear thanks….

  Thank you for the laughs, the tears, the screams

  Gratitude for the late nights, the dancing, the dreams

  Thank you for the walks, the talks, and meals together

  Many thanks for the shenanigans I’ll cherish forever

  Thank you for the compassion, for the coffee and tea

  Appreciation for the love, the hate, and learning to be

  Thank you for the support, and most ardently, for embracing me

bibliography: bib/thesis.bib
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line).

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

## The historical significance of variability 
It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bouchard 2014, Bambach 1993). Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Soo et al 2017, Blankenship 2017). Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll 2008). In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnik et al 2007). Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Reed et al 2010, Burggren 2018). Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.

## Relevant significance of variability today 
In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e. El Nino Southern Oscillation).  As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (Khelifa et al 2019, Stewart et al 2013). How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e. hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Helmuth et al 2010, Bernardo 2014, Huey et al 2012).

## Why this study on variability matters 
There is a large body of literature on environmental variability including its relevance for positive effects for populations and communities, i.e. inflationary effects in sink-source systems (Holt et al 2003, Vasseur and Fox 2009), such that variability aids population density when organisms emigrate. However, there is a lack of consensus as to which scales of biological organization, i.e. individuals, populations, or communities of organisms will be most severely affected by different shifts in environmental conditions (problem statement). We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response (aims). To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.