### 2021-04-29 writing /thesis_draft/05-appendix.Rmd

`r if(knitr:::is_latex_output()) '\\appendix'`

# Appendix {-}

```{r, echo=FALSE, results = 'hide', message=FALSE}
pacman::p_load(metafor, tidyverse, viridis, visreg, forcats, kableExtra)
dat_MA_0<- read_csv("metafor.csv") %>%
  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))

#make standardized SD columns based on N and SE vs SD
dat_MA <-dat_MA_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_MA_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_MA, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) 
count(dat_MA_ES, study_id)

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -4) 
```

```{r, echo = FALSE}
#funnel plot 
### set up 2x2 array for plotting
par(mfrow=c(2,2))
funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
```

```{r, echo=FALSE, message = FALSE, fig.cap="Funnel plot displaying the extracted data points and the distribution of their residuals centered around zero.", fig.align="center", out.width = '90%', fig.scap="Funnel plot"}
### draw funnel plots
#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
```

\clearpage

```{r, echo= FALSE, fig.cap="Distribution of effect sizes (n=140) across data from all 15 studies.", fig.align="center", out.width = '90%', fig.scap="Histogram of effect sizes"}
dat_MA_ES %>%
filter(yi > -20) %>%
  ggplot(aes(x=yi))+
  geom_histogram(binwidth = 1)+
  theme_bw()+
  scale_fill_viridis()
```

```{r, echo= FALSE}
table4 <- data.frame(Species = c("Drosophila melanogaster", "Leptopilina boulardi", 
                         "Mauremys reevesii", "Gryllus firmus", 
                         "Limnodynastes peronii", "Limnodynastes tasmaniensis", 
                         "Platyplectrum ornatum ", "Natrix natrix", 
                         "Coturnix japoncia", "Plestiodon chinensis", 
                         "Rhodnis prolixus", "Escherichia coli", 
                         "Salmonella", "Caiman latirostris", 
                         "Trachemys scripta", "Esteya vermicola"), 
                 Topt = c("24-26.6°C", "25°C", "37.97 ± 0.64 °C", "~30°C", "30°C", "30-33°C",
                          "33°C", "~30°C", "26°C", "33°C", "30-35°C", "37°C, 35-36°C",
                          "35-36°C", "32-33°C", "32-33°C", "26°C"), 
                 `Reference` = c("Klepsatel 2013, David 1988", "De lava et al 2016", 
                                 "Dang et al 2019", "Singh et al 2020", "Seebacher et al 2014",
                                 "Whitehead et al 1989", "Kern et al 2014", "Issac 1997",
                                 "Alagawany et al 2017", "Baojun et al 2010", "Fresquet and Lazzari 2011",
                                 "Doyle and Schoeni 1984, Bronikowski et al 2001", 
                                 "Bronikowski et al 2001", 
                                 "Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002", 
                                 "Dang et al 2019", "Wang et al 2015"))
kable(table4, "latex", booktabs = T, 
      col.names = c("Species", "Topt (°C)", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X. arvicola, used in Garcia-Ruiz et al 2011)", caption.short ="Thermal stress data by species") %>%
   row_spec(0,bold=TRUE) %>%
  column_spec(3, width = "6cm") %>%
   kable_styling(latex_options = c("striped", "hold_position"))
```