### 2021-04-30 writing /thesis_draft/03-chap3.Rmd

```{r include_packages_2, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis and also two functions
# used for labeling and referencing
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("remotes")',
        "first in the Console."
      )
    )
  }
}
if (!require(dplyr)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("dplyr", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("dplyr")',
        "first in the Console."
      )
    )
  }
}
if (!require(ggplot2)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("ggplot2", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("ggplot2")',
        "first in the Console."
      )
    )
  }
}
if (!require(bookdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("bookdown", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("bookdown")',
        "first in the Console."
      )
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
flights <- read.csv("data/flights.csv", stringsAsFactors = FALSE)
```

# Results {}

```{r, echo=FALSE, results = 'hide', message=FALSE}
pacman::p_load(metafor, tidyverse, viridis, visreg, forcats, kableExtra)
dat_MA_0<- read_csv("metafor.csv") %>%
  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))

#make standardized SD columns based on N and SE vs SD
dat_MA <-dat_MA_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_MA_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_MA, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) 
count(dat_MA_ES, study_id)

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -4) 
```

```{r, echo= FALSE, message= FALSE, fig.cap="Forest plot of effects sizes by study and experiment within a study. Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study. Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", fig.scap="Forest plot of effect sizes", out.width="95%", fig.align="center"}
forest <- dat_MA_ES %>%
  group_by(study_id, experiment_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2", 
       header=TRUE, cex=0.77)
```

Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 3.1. These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1). As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.

Our full model, including 7 covariates (all, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001). Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p < 0.00001). Organization level (n=140) was also important in the model, as a statistically significant covariate (df= 7, p < 0.05).

```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficient estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.  Error bars represent 95 percent confidence intervals.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML")

#plot to include to demonstrate model estimates
full_model_coefficients<- as.data.frame(coef(full_rf_model)) %>%
  tibble::rownames_to_column("covariates") %>%
  rename(value = "coef(full_rf_model)") %>%
  arrange(desc(value)) %>%
  mutate(coefficients = c("intercept", "response trait type", 
                          "fluctuation range", "mean temperature",
                          "fluctuation range:mean temperature",
                                       "size", "experimental age","organization level" 
                                      ), 
         lb = c(-0.9781, -0.1371, 0.1794, -0.0184, -0.0177, -0.8898, -1.3401, -1.2954), 
         ub= c(2.7433, 1.3568, 0.3711, 0.0536, -0.0097, 0.2348, 0.2367, 0.0041))

label.df <- data.frame(coefficients = c("fluctuation range", "fluctuation range:mean temperature", "organization level"),
                       value = c(0.5, -0.25, -1.4))

ggplot(data=full_model_coefficients, aes(x = reorder(coefficients, value),
             y = value,
             fill = value > 0)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), width=.2,
                 position=position_dodge(.95), color = "black")+ 
  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black") +
  labs(x = NULL, y = "Coefficient estimate") +
  scale_fill_viridis(discrete = TRUE)+
  theme_bw()+
  coord_flip()+
  geom_text(data=label.df, label = "*", color = "black", size = 6)
```

\clearpage

Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2). The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p < 0.05. Surprisingly, the additional negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant. In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).

```{r, echo= FALSE}
table3 <- data.frame(Cat = c("intercept", "stressful"),
                     estimate = c(0.2050, -0.0489), 
                 se = c(0.1846, 0.0702),
                 zval = c(1.1104, -0.6975),
                 pval = c(0.2668, 0.4855), 
                 ci.lb = c(-0.1568, -0.1864), 
                 ci.ub = c(0.5669, 0.0886))
kable(table3, "latex", booktabs = T, col.names = c("Term", "estimate", "se", "zval", "pval", "ci.lb", "ci.ub"), escape = F, caption = "Model summary statistics for random effects model using thermal stress as sole modifier", 
      caption.short ="Thermal stress model summary statistics") %>%
   row_spec(0,bold=TRUE)%>%
   kable_styling(position = "center", latex_options = c("striped", "hold_position"))
```

\clearpage

The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3. At lower mean temperatures, higher fluctuation ranges have a more positive effect on organism responses and performance, though that trend starkly ends at about $24^{\circ}$C. Beyond $24^{\circ}$C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).

```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and effect size. Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C. Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure 2).", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and effect size"}
normalized %>%
ggplot(aes(x=mean_temp_constant, y = yi, fill=flux_range))+
  geom_jitter( alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Temperature range (°C)', option = "C")+
  theme_classic()+
  xlab("Mean Temperature (°C)")+
  ylab("Effect Size")+
  geom_line(y=0, linetype = "dashed")+
  theme(legend.position="bottom")
```

The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4). Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).

As noted in the methods, when we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p < 0.001) as did mean temperature (df=8, p < 0.001). Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p < 0.1).

```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by organization level: individual or population level responses reported in studies.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by organization level"}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(org_level)))+
  geom_point()+
  labs(x = "Temperature fluctuation range (°C)",
       y ="SMD",
       color = "Biological organization level")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("individual", "population"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom") 
```

\clearpage

```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by thermal stress.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by thermal stress"}
normalized %>%
  drop_na(stressful) %>%
ggplot(aes(x=flux_range, y=yi, color = factor(stressful)))+
  geom_jitter()+
  labs(x = "Temperature fluctuation range (°C)",
       y ="SMD",
       color = "Thermally stressfull")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("no", "yes"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by life history metric: responses categorized as rate or trait.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by response type"}
normalized %>%
ggplot(aes(x=flux_range, y=yi, color = resp_type))+
  geom_jitter()+
  labs(x = "Temperature fluctuation range (°C)",
       y ="SMD",
       color = "Life history metric")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("rate", "trait"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by body size: extra-small, small, medium, or large.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by body size"}
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(size)))+
  geom_point()+
  labs(x = "Temperature fluctuation range (°C)",
       y ="SMD",
       color = "Size")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("extra-small", "small", "medium", "large"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by experimental age: larval, juvenile, or adult.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by experimental age"}
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(exp_age)))+
  geom_point()+
  labs(x = "Temperature fluctuation range (°C)",
       y ="SMD",
       color = "Experimental age")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("larval", "juvenile", "adult"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

# Discussion {}

The results of this meta-analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level. There were not overall differences in the direction of thermally variable environments positively or negatively impacting performance relative to constant environments. However, a considerable amount of variation exists in whether thermally variably environments positively, negatively, or minimally impact performance across studies. Our model results highlight the importance of fluctuation range, mean temperature, and organization level to better understand the variety of effects thermal variation has on organismal responses. These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e. higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures. The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change [@vasseur_increased_2014; @sunday_thermal_2012]. With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed “the point of no return” [@russill_tipping_2009], understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades [@west_allometric_2002; @cheung_application_2008]. In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature [@vasseur_increased_2014].

These findings underscore the significance of better understanding why differences across studies and experimental designs occurred. Jensen’s inequality is one way to conceptualize why responses differ, as averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance [@ruel_jensens_1999; @bernhardt_nonlinear_2018]. The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive. Acclimation may also explain differences in response across studies as the scale on which variation occurs influences how organisms may respond. If organisms’ non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (i.e. inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies [@kremer_gradual_2018]. In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance. Rapid evolution could further explain differences in responses across studies. If organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics [@yoshida_rapid_2003]. It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.

Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal stress, life history traits, and experimental age. Body size has been identified as important in the allometric scaling of MTE and it is interesting that in these subsets of studies, body size did not explain the differences in responses across studies. This could simply be because we had a small sample size of studies (n=15), though multiple responses within each study contributed to a large sample of effect sizes (n=140), there were only a subset of relative body sizes included. It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms [@clarke_scaling_1999; @clarke_temperature_2006]. This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns that variability in temperature drives.

An important side note is that our full model and thermal stress model differed in the number of studies, as one study, @garcia-ruiz_effects_2011, did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola). Further, this study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding @garcia-ruiz_effects_2011 from our full model. We decided to include this study in our analysis because data were extracted from a table, therefore there were no inaccuracies in obtaining the data. It provides an important counter and discussion for how and why responses from different studies may be so different from others.

While body size and thermal stress emerging as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship between thermal variability and MTE. If we were able to obtain additional information about the CTmax, CTmin, thermal breadth, and tolerance range for each of the species included in the analysis, we may also be able to better understand the trends and the significance of variability patterns in dictating responses.

Both the type of response (i.e. whether it was a rate or trait) and experimental age were also not important predictors in our full model. We expected rates and traits to have different effects, as traits may be more heavily affected by variability because they are not as temporally dynamic as rates. The lack of difference between rates and traits may be because variability is not differentially acting on rate or trait based responses, simply dampening their overall effects. Again, this may be an artifact of having a small sample size, with mainly trait based responses as opposed to rate based responses.

We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018). This may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile stage instead. However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018]. These results align with patterns of juvenile organisms (i.e. fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance [@portner_ecology_2008; @moyano_effects_2017].

Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project. As Figure 1.3 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures. It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated. Are these diurnal temperature fluctuations accurate for what is expected under natural conditions? This lack of variety in pattern may be due to a lack of consensus on the language used to classify and discuss thermal variability. Though there is a large body of literature that investigates temperature variation, there appears to be dissonance in the scientific literature about the language used to discuss such variation. Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures [@amarasekare_framework_2012] as well as different ranges of temperature, both of which are different concepts. For instance, @amarasekare_framework_2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits. Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern. Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes [@colinet_mechanisms_2018; @colinet_insects_2015]. Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.

Many of these studies failed to justify why certain temperatures were chosen as mean temperatures as well as the range of temperatures studied. In order to compile information on the thermal stress of each of the organisms included in the studies of this meta-analysis, additional resources were needed (see Appendix Table 1). Understanding where these organisms' performance falls in relation to temperature on its TPC is crucial for drawing reliable conclusions of how thermal variability affects performance. Without it, only broad conclusions can be drawn, as have been drawn here, that variability and range affects thermal performance.

Several studies made little to no mention of generation time with respect to duration of their experiments or study organisms used. How organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020]. This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis. Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times. The converse is true for seasonal or decadal thermal fluctuations.

The results of this analysis differ from previous attempts to understand variability as they quantitatively address how variability affects organisms across taxa, not simply one group. Understanding how allometric scaling and thermal variability patterns interact is important for predicting realistic organismal responses in the face of variable conditions in the environment.